### 2016-05-01 writing/latex/wilf.tex
Sentence ordering and autovars

The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
Wilf-class since the only permutation containing this pattern is
the permutation \(\perm{1}\). The avoiders of this pattern therefore
have generating function \(\var(x) = C(x)-x\).\nextvar

The avoiders of the pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) can be
decomposed around the maximum element in order to give the following
structural decomposition.
\begin{equation*}
    \mathcal{\var} = \varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{\var}}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p_1\). The lower right region however can create occurrences
of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

@@

It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(p_2 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of an avoider of \(p_2\) in
\(\av{\perm{2,3,1}}\) around the maximum\nextvar[\varullr]
\begin{equation*}
    \mathcal{\var}_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(p_2\) starts with it's maximum, by not allowing the left part of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.

@@

There is one pattern of length \(1\) still to consider. The pattern
\(p_5 = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(p_5\) must fit into the following form precisely once.
\begin{equation*}
    \mathcal{H} = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{H}}}_\text{
        Minimum is after the maximum, cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(H(x)\) satisfying
\begin{equation*}
    H(x) = 1 + xC(x)(C(x)-1) + x(H(x)-1)
\end{equation*}
This generating function can be enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\end{equation*}

@@

\subsection{}
The set/symmetry classes containing the following patterns are
Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not set equivalent or
symmetries of each other
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/2,
                    1/0,1/1,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/0,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
In this case it is better to consider the containers of the patterns
instead of the avoiders due to the amount of shadings in the mesh.

@@

Consider an avoider of \(\perm{2,3,1}\) and \(m_3\). We can perform
the decomposition around the maximum
\begin{equation*}
    \mathcal{I}_1 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\mathcal{I}_1}{}{}{\mathcal{G}_1}
\end{equation*}
Since only the first point in the top right region can
create an occurrence of \(m_3\) if and only if it is the
maximum in this region we must avoid starting with the maximum.

Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
can perform the same decomposition around the maximum to get
\begin{equation*}
    \mathcal{I}_2 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\mathcal{I}_2}{}{}{\mathcal{G}_3}
\end{equation*}
Since an occurrence of \(m_4\) can never occur in the
top right region, and could only occur between the first point
in the region and the maximum, if and only if this first point is the minimum.
Since both \(\mathcal{G}_1\) and \(\mathcal{G}_3\) have the same
enumeration, \(\mathcal{I}_1\) and \(\mathcal{I}_2\) must also have
the same enumeration and are therefore Wilf-equivalent.

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\perm{2,3,1},m_1}\) we gain the following.
\begin{equation*}
    \mathcal{I}_3 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\mathcal{I}_3}{}{}{\mathcal{G}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore all four patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\) with generating function satisfying
\begin{equation*}
    I(x) = 1 + xI(x)G(x)
\end{equation*}
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc
\end{equation*}
This corresponds to OEIS sequence number \href{https://oeis.org/A035929}{A035929}
with offset 1.

\subsection{}
The classes containing the following patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
It is obvious that these two are Wilf-equivalent since the only
permutations that contain these patterns are \(\perm{1,2}\) and
\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
are counted by the Catalan numbers at all lengths except for length
\(2\) where there is precisely \(1\) avoider.
Therefore there are
\begin{equation*}
    1, 1, 1, 5, 14, 42, 132, 429, 1430, 4862, 16796,\cdots
\end{equation*}

@@

First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\)
we can perform the usual structural decomposition of an avoider
of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
precisely once.
\begin{equation*}
    \mathcal{J}_1 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{F}}
\end{equation*}

Now consider the decomposition around the maximum of a permutation in
\(\av{\perm{2,3,1},m_2}\) This fits into the following set.
\begin{equation*}
    \mathcal{J}_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
    \decompmax{0/1,1/0,1/1}{\mathcal{F}}{}{}{}
\end{equation*}
Therefore both of these sets of avoiders are enumerated in the same manner
having generating function satisfying
\begin{equation*}
    J(x) = 1 + xC(x)(C(x)-1) + xF(x)
\end{equation*}
This generating function gives
\begin{equation*}
    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc
\end{equation*}

\subsection{}
Consider the containers of the patterns
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
\end{equation*}

For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the free box in each case. Each of these regions
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    K_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
\end{equation*}

\subsection{}
Now consider the containers of the patterns
\begin{equation*}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Each of these patterns again occurs precisely once in
any containing permutation. However this time when considering
the free box we need to take into consideration \LemmaRef{lem:incdec}
and so the empty box can only contain a decreasing subsequence. There
is precisely one decreasing subsequence at every length, and so there
is exactly one container of each pattern at each length. The three
patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
length \(n\) for all \(n \ge 2\).
This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
\end{equation*}

@@

First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
around the maximum.
\begin{equation*}
    \mathcal{L}_1 =
    \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle \mathcal{G}_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{L}_1}
\end{equation*}

Now we decompose the avoiders of \(m_2\) around the leftmost point
\begin{equation*}
    \mathcal{L}_2 =
    \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
    \sqcup
    \decompleft{0/1,1/0,1/1}{\mathcal{L}_2}{}{}{}
\end{equation*}
This gives us two generating functions satisfying the following pair of equations
\begin{align}
    L_1(x) &= 1 + xC(x)(G(x)-1) +xL_1(x) \label{eqn:genL1}\\
    \text{ and }L_2(x) &= 1 + x(C(x)-1)^2 + xL_2(x) \label{eqn:genL2}
\end{align}
In order for these two functions to give the same value it is necessary to show that
\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
\(C(x)(G(x)-1) = (C(x)-1)^2\).
\begin{align*}
    &&(C(x)-1)^2 &= C(x)(G(x)-1)\\
        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(G\) and \(C\)}\\
        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
\end{align*}
The final line is always satisfied since it is the form of \(C(x)\), and
therefore the two generating functions are equal.

@@

Evaluating the generating function \(L(x)\) gives us the enumeration
\begin{equation*}
    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc
\end{equation*}

\subsection{}
The classes containing the following patterns are Wilf-equivalent
\begin{equation*}
    \pattern{}{1,2}{} \text{ and }
    \pattern{}{2,1}{}
\end{equation*}
This is true since the only avoiders of these patterns are the decreasing
sequence and the increasing sequence respectively, and both of these
avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
every length.

\subsection{}
The patterns
\begin{equation*}
    \pattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and }
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/1,1/2}
\end{equation*}
are Wilf-equivalent.
If \(\mathcal{M}_1\) is the set of avoiders of
\textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
decomposition around the maximum we have
\begin{equation*}
    \mathcal{M}_1 = \varepsilon \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
\end{equation*}
This is because \(\mathcal{M}_1\) is the set of permutations
who have their minimum occur after their maximum.

% If \(\mathcal{M}_2\) is the set of avoiders of
% \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2}, then by structural
% decomposition around the maximum we have
% \begin{equation*}
%     \mathcal{M}_2 = \varepsilon \sqcup
%     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{M}_1\setminus\varepsilon}
%     \sqcup \tikz{\filldraw (0,0) circle (3pt);}
% \end{equation*}
% Alternatively,
The pattern \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2} occurs if
the last element is higher than the penultimate element. This can only occur
if the last element is in a single block in the set partition context, In order
to construct a avoider of length \(n\) we can take any avoider of
\(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
block.  This ensures that the last block is never a singleton. This means
that these permutations are also counted by \(M(x) = 1+xC(x)\).

The avoiders of the third pattern can be decomposed by the maximum
to give
\begin{equation*}
    \mathcal{M}_3 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{F}}{}{}{\mathcal{M}_3}
\end{equation*}
The generating function derived satisfies \(M_3(x) = 1 + xF(x)M_3(x)\).
The fourth pattern can be decomposed around the maximum in a similar
manner.
\begin{equation*}
    \mathcal{M}_4 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{M}_4}{}{}{\mathcal{F}}
\end{equation*}

Finally considering the last pattern, the only way we can construct an
avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
at the start of the permutation. Adding a new leftmost point in any other
position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
Therefore this is also counted by \(M(x) = 1+xC(x)\).

We need to show that the generating function \(M_3(x)\) is the same as \(M(x)\)
\begin{equation*}
    \begin{aligned}
        M_3(x) &= 1 + xF(x)M_3(x) \\
        &= \frac{1}{1-xF(x)} && \text{Solving for \(M_3(x)\)}\\
        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(F(x)\)}\\
        &= \frac{1-xC(x) +x}{1-xC(x)}\\
        &= 1+xC(x)
    \end{aligned}
\end{equation*}
We have that \(M_3(x) = 1+xC(x)= M(x) \) so all four patterns are Wilf-equivalent
 and have enumeration sequence
 \begin{equation*}
     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc
 \end{equation*}
 This is an offset of the Catalan numbers.

>>>>

The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
Wilf-class since the only permutation containing this pattern is
the permutation \(\perm{1}\). The avoiders of this pattern therefore
have generating function \(\var(x) = C(x)-x\).

\nextvar
The avoiders of the pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) can be
decomposed around the maximum element in order to give the following
structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p_1\). The lower right region however can create occurrences
of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

@@

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(p_2 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of an avoider of \(p_2\) in
\(\av{\perm{2,3,1}}\) around the maximum
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(p_2\) starts with it's maximum, by not allowing the left part of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.

@@

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(p_5 = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(p_5\) must fit into the following form precisely once.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function can be enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\end{equation*}

@@

\subsection{}
\nextvar[\varminl]
The set/symmetry classes containing the following patterns are
Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not set equivalent or
symmetries of each other
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/2,
                    1/0,1/1,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/0,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
In this case it is better to consider the containers of the patterns
instead of the avoiders due to the amount of shadings in the mesh.

@@

Consider an avoider of \(\perm{2,3,1}\) and \(m_3\). We can perform
the decomposition around the maximum
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_1}{}{}{\mathcal{\varmaxl}_1}
\end{equation*}
Since only the first point in the top right region can
create an occurrence of \(m_3\) if and only if it is the
maximum in this region we must avoid starting with the maximum.

Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
can perform the same decomposition around the maximum to get
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{1/0,0/1}{\scriptvar_2}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
Since an occurrence of \(m_4\) can never occur in the
top right region, and could only occur between the first point
in the region and the maximum, if and only if this first point is the minimum.
Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
enumeration, \(\scriptvar_1\) and \(\scriptvar_2\) must also have
the same enumeration and are therefore Wilf-equivalent.

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\perm{2,3,1},m_1}\) we gain the following.
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore all four patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc
\end{equation*}
This corresponds to OEIS sequence number \href{https://oeis.org/A035929}{A035929}
with offset 1.

\subsection{}
\nextvar
The classes containing the following patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
It is obvious that these two are Wilf-equivalent since the only
permutations that contain these patterns are \(\perm{1,2}\) and
\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
are counted by the Catalan numbers at all lengths except for length
\(2\) where there is precisely \(1\) avoider.
Therefore there are
\begin{equation*}
    1, 1, 1, 5, 14, 42, 132, 429, 1430, 4862, 16796,\cdots
\end{equation*}

@@

First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\)
we can perform the usual structural decomposition of an avoider
of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
precisely once.
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}

Now consider the decomposition around the maximum of a permutation in
\(\av{\perm{2,3,1},m_2}\) This fits into the following set.
\begin{equation*}
    \scriptvar_2 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
    \decompmax{0/1,1/0,1/1}{\mathcal{\varfine}}{}{}{}
\end{equation*}
Therefore both of these sets of avoiders are enumerated in the same manner
having generating function satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x\varfine(x)
\end{equation*}
This generating function gives
\begin{equation*}
    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc
\end{equation*}

\subsection{}
\nextvar
Consider the containers of the patterns
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
\end{equation*}

For each of these patterns there is precisely one occurrence
in any permutation containing the pattern. Now consider
the points in the free box in each case. Each of these regions
must contain an avoider of \(\perm{2,3,1}\) that is of length
\(n-2\). Therefore these classes are all Wilf-equivalent
and the number of length \(n\) avoiders is
\begin{equation*}
    \var_n = C_n - C_{n-2}
\end{equation*}
for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
\end{equation*}

\subsection{}
\nextvar
Now consider the containers of the patterns
\begin{equation*}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Each of these patterns again occurs precisely once in
any containing permutation. However this time when considering
the free box we need to take into consideration \LemmaRef{lem:incdec}
and so the empty box can only contain a decreasing subsequence. There
is precisely one decreasing subsequence at every length, and so there
is exactly one container of each pattern at each length. The three
patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
length \(n\) for all \(n \ge 2\).
This gives the sequence
\begin{equation*}
    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
\end{equation*}

@@

First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
around the maximum.
\begin{equation*}
    \scriptvar_1 =
    \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
\end{equation*}

Now we decompose the avoiders of \(m_2\) around the leftmost point
\begin{equation*}
    \scriptvar_2 =
    \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
    \sqcup
    \decompleft{0/1,1/0,1/1}{\scriptvar_2}{}{}{}
\end{equation*}
This gives us two generating functions satisfying the following pair of equations
\begin{align}
    \var_1(x) &= 1 + xC(x)(\varmaxl(x)-1) +x\var_1(x) \label{eqn:genL1}\\
    \text{ and }\var_2(x) &= 1 + x(C(x)-1)^2 + x\var_2(x) \label{eqn:genL2}
\end{align}
In order for these two functions to give the same value it is necessary to show that
\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
\(C(x)(\varmaxl(x)-1) = (C(x)-1)^2\).
\begin{align*}
    &&(C(x)-1)^2 &= C(x)(\varmaxl(x)-1)\\
        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(\varmaxl\) and \(C\)}\\
        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
\end{align*}
The final line is always satisfied since it is the form of \(C(x)\), and
therefore the two generating functions are equal.

@@

Evaluating the generating function \(\var(x)\) gives us the enumeration
\begin{equation*}
    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc
\end{equation*}

\subsection{}
\nextvar
The classes containing the following patterns are Wilf-equivalent
\begin{equation*}
    \pattern{}{1,2}{} \text{ and }
    \pattern{}{2,1}{}
\end{equation*}
This is true since the only avoiders of these patterns are the decreasing
sequence and the increasing sequence respectively, and both of these
avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
every length.

\subsection{}
The patterns
\begin{equation*}
    \pattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,1/0,1/2,2/0,2/2},
    \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and }
    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/1,1/2}
\end{equation*}
are Wilf-equivalent.
If \(\scriptvar_1\) is the set of avoiders of
\textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
decomposition around the maximum we have
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
\end{equation*}
This is because \(\scriptvar_1\) is the set of permutations
who have their minimum occur after their maximum.

% If \(\mathcal{M}_2\) is the set of avoiders of
% \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2}, then by structural
% decomposition around the maximum we have
% \begin{equation*}
%     \mathcal{M}_2 = \varepsilon \sqcup
%     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{M}_1\setminus\varepsilon}
%     \sqcup \tikz{\filldraw (0,0) circle (3pt);}
% \end{equation*}
% Alternatively,
The pattern \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2} occurs if
the last element is higher than the penultimate element. This can only occur
if the last element is in a single block in the set partition context, In order
to construct a avoider of length \(n\) we can take any avoider of
\(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
block.  This ensures that the last block is never a singleton. This means
that these permutations are also counted by \(\var(x) = 1+xC(x)\).

The avoiders of the third pattern can be decomposed by the maximum
to give
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{\varfine}}{}{}{\scriptvar_3}
\end{equation*}
The generating function derived satisfies \(\var_3(x) = 1 + x\varfine(x)\var_3(x)\).
The fourth pattern can be decomposed around the maximum in a similar
manner.
\begin{equation*}
    \scriptvar_4 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptvar_4}{}{}{\mathcal{\varfine}}
\end{equation*}

Finally considering the last pattern, the only way we can construct an
avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
at the start of the permutation. Adding a new leftmost point in any other
position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
Therefore this is also counted by \(\var(x) = 1+xC(x)\).

We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
\begin{equation*}
    \begin{aligned}
        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
        &= \frac{1-xC(x) +x}{1-xC(x)}\\
        &= 1+xC(x)
    \end{aligned}
\end{equation*}
We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
 and have enumeration sequence
 \begin{equation*}
     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc
 \end{equation*}
 This is an offset of the Catalan numbers.
